---
title: "Tidy Tuesday: Internet Access"
date: '`r format(Sys.time(), "%d %B, %Y")`'
author: "Nozomi Nakajima"
output: html_document
editor_options: 
  chunk_output_type: console
---

The data this week comes from [Microsoft](https://github.com/microsoft/USBroadbandUsagePercentages) by way of [The Verge](https://www.theverge.com/22418074/broadband-gap-america-map-county-microsoft-data).

> If broadband access was a problem before 2020, the pandemic turned it into a crisis. As everyday businesses moved online, city council meetings or court proceedings became near-inaccessible to anyone whose connection couldn’t support a Zoom call. Some school districts started providing Wi-Fi hotspots to students without a reliable home connection. In other districts, kids set up in McDonald’s parking lots just to get a reliable enough signal to do their homework. After years of slowly widening, the broadband gap became impossible to ignore.

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

# load libraries
library(tidyverse)
library(scales)
library(countrycode)
library(ggthemes)
library(ggmap)
library(gganimate)
library(zipcodeR)

# set theme
theme_set(theme_light())

# set working directory
setwd("~/Documents/GitHub/TidyTuesday/2021/20210511_InternetAccess")

```


Load weekly data:

```{r load, echo = FALSE}
broadband <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-11/broadband.csv')

```

Data exploration:

```{r, include = TRUE}
 
```

