---
title: "Tidy Tuesday: Internet Access"
date: '`r format(Sys.time(), "%d %B, %Y")`'
author: "Nozomi Nakajima"
output: html_document
editor_options: 
  chunk_output_type: console
---

The data this week comes from [Microsoft](https://github.com/microsoft/USBroadbandUsagePercentages) by way of [The Verge](https://www.theverge.com/22418074/broadband-gap-america-map-county-microsoft-data).

> If broadband access was a problem before 2020, the pandemic turned it into a crisis. As everyday businesses moved online, city council meetings or court proceedings became near-inaccessible to anyone whose connection couldn’t support a Zoom call. Some school districts started providing Wi-Fi hotspots to students without a reliable home connection. In other districts, kids set up in McDonald’s parking lots just to get a reliable enough signal to do their homework. After years of slowly widening, the broadband gap became impossible to ignore.

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

# load libraries
library(tidyr)
library(dplyr)
library(tidyverse)
library(stringr)
library(scales)
library(countrycode)
library(ggthemes)
library(ggmap)
library(gganimate)
library(zipcodeR)
library(tigris)
library(tidycensus)
library(readxl)
library(janitor)

# set theme
theme_set(theme_light())

# set working directory
setwd("~/Documents/GitHub/TidyTuesday/2021/20210511_InternetAccess")

```


Load weekly data:

```{r load, echo = FALSE}
bb <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-11/broadband.csv') %>% 
  janitor::clean_names() %>% 
    ## makes all lower case
  rename(state = st) %>%
  mutate(state = state.name[match(state, state.abb)]) %>%
          ## R has state.name and state.abb vectors
  mutate(state = ifelse(is.na(state), "District of Columbia", state),
          ## if missing state, then DC, otherwise keep state 
         broadband_availability_per_fcc = parse_number(broadband_availability_per_fcc, na = "-"),
          ## change to numeric broadband_availability_per_fcc
          ## some missing is coded with - so replace those as na
         broadband_usage = parse_number(broadband_usage, na = "-")) %>%
    mutate(county = paste0(str_remove(county_name, " County$"), ", ", state),
          ## new var county removes the word "county" at the end of each label
          ## add , and state name to the label 
         county = fct_reorder(county, broadband_availability_per_fcc),
          ## reorder the new var county by population size
         county_id = sprintf("%05d", county_id))
          ## fix this so it matches with county_sf shape file
          
bbz <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-11/broadband_zip.csv') %>%
  janitor::clean_names() %>% 
  rename(state = st) %>%
  mutate(state = state.name[match(state, state.abb)]) %>%
  mutate(state = ifelse(is.na(state), "District of Columbia", state)) %>%
  mutate(county = paste0(str_remove(county_name, " County$"), ", ", state),
         postal_code = sprintf("%05d", postal_code))
          ## sprintf formats number with fixed width; specify adding leading 0s to ensure 5 digits
  
```


```{r load, echo = FALSE}

# can take census data using tidycensus (but this  needs api)

# read in using downloaded csv file from: https://www.census.gov/data/datasets/time-series/demo/popest/2010s-counties-total.html

county_pop <- read_excel("~/Documents/GitHub/TidyTuesday/2021/20210511_InternetAccess/co-est2019-annres.xlsx", skip = 3) %>%
  select(county = ...1, population_2017 = '2017',
         population_2019 = '2019') %>% 
  separate(county, c("county_name", "state"), sep = ", ") %>%
  mutate(county_name = str_remove(county_name, "^\\."))
  ## start of string is ^ regular expression is \\

# check merge
bb %>% 
  left_join(county_pop, by = c("county_name", "state")) %>%
  filter(is.na(population_2017)) 
  # four counties with missing population in 2017

# merge two data sets using inner_join
bb_with_pop <- bb %>% 
  inner_join(county_pop, by = c("county_name", "state")) 

# let's look at 25 most populated counties
bb_with_pop %>%
  arrange(desc(population_2017)) %>% 
  head(25) %>%
  mutate(county = fct_reorder(county, population_2017)) %>%
  ggplot(aes(population_2017, county)) +
  geom_col()
  
# let's look at bb availability in 25 most populated counties
bb_with_pop %>%
  arrange(desc(population_2017)) %>% 
  head(40) %>%
  mutate(county = fct_reorder(county, broadband_availability_per_fcc)) %>%
    ## order by bb availability
  ggplot(aes(broadband_availability_per_fcc, county)) +
  geom_point() +
  scale_x_continuous(labels = percent_format())
    ## display x axis as a percent

# relationship between population and broadband availability
bb_with_pop %>%
  filter(population_2017 >= 30000 ) %>%
    ## filter to larger counties
  ggplot(aes(population_2017, broadband_availability_per_fcc)) +
  geom_point() +
  geom_text(aes(label = county), 
            check_overlap = TRUE, vjust = 1, hjust = 1) +
              # only first plot will be plotted (in order of bb availability, then pop)
  scale_x_log10(labels = comma_format()) +
    ## display x axis population in log scale
  scale_y_continuous(labels = percent_format()) +
    ## display y axis as a percent  
  expand_limits(x = 10000) +
    ## moves the x-axis to start at 10,000 so we can see the county labels
  labs(y = "Broadband availability in 2017",
       x = "Population in 2017")

# relationship between population and broadband usage
bb_with_pop %>%
  filter(population_2019 >= 1000 ) %>%
    ## filter to large counties
  ggplot(aes(population_2019, broadband_usage)) +
  geom_point() +
  geom_text(aes(label = county), 
            check_overlap = TRUE, vjust = 1, hjust = 1) +
              # only first plot will be plotted (in order of bb availability, then pop)
  geom_smooth(method = "lm") +
  scale_x_log10(labels = comma_format()) +
    ## display x axis population in log scale
  scale_y_continuous(labels = percent_format()) +
    ## display y axis as a percent  
  expand_limits(x = 10000) +
    ## moves the x-axis to start at 10,000 so we can see the county labels
  labs(y = "Broadband usage in 2019",
       x = "Population in 2019")


```


Zip code data: 

```{r, include = TRUE}

# from the zipcode R package: 
as_tibble(zip_code_db)


# merge bbz with zipcode file
bbz_zip_join <- bbz %>% 
  inner_join(zip_code_db %>% 
               select(-state, -county), ## filter out state and country from zip_code_db so we use the ones from bbz
             by = c(postal_code = "zipcode") )
            ## var joining by is postal_code in bbz and zipcode in zip_code_db


# broadband usage by population density 
bbz_zip_join %>%
  filter(population >= 10000) %>%
    ## only look at zipcodes with at least 10,000 people
  ggplot(aes(population_density, broadband_usage)) +
  geom_point() +
  scale_x_log10(labels = comma_format()) +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Persons per square mile",
       y = "Broadband usage in 2019")

# is population density bimodal? a bit.
bbz_zip_join %>%
  ggplot(aes(population_density)) +
  geom_histogram() +
  scale_x_log10()

# plot for biggest states
bbz_zip_join %>%
  count(state, sort = T)

# plot for largest states 
bbz_zip_join %>%
  filter(population >= 10000) %>%
  filter(state %in% c("Texas", "Pennsylvania", "New York", "California")) %>%
  ggplot(aes(population_density, broadband_usage)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ state) + 
  scale_x_log10(labels = comma_format()) +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Persons per square mile",
       y = "Broadband usage in 2019")

```


Use tigris package to get shapefile

```{r, include = TRUE}
library(tigris)
library(sf)

# get us counties shapefile
?counties()

counties_sf <- counties() %>%
  st_simplify(dTolerance = .03) 

counties_sf %>%
  inner_join(bb_with_pop, by = c(GEOID = "county_id")) %>%
  filter(!state %in% c("Alaska", "Hawaii")) %>%
  ggplot() + 
  geom_sf() +
  theme_map() +
  coord_sf()

counties_sf %>%
  inner_join(bb_with_pop, by = c(GEOID = "county_id")) %>%
  filter(!state %in% c("Alaska", "Hawaii")) %>%
  ggplot(aes(fill = broadband_availability_per_fcc)) + 
  geom_sf() +
  scale_fill_continuous(labels = percent) +
  theme_map() +
  coord_sf()

counties_sf %>%
  inner_join(bb_with_pop, by = c(GEOID = "county_id")) %>%
  filter(!state %in% c("Alaska", "Hawaii")) %>%
  ggplot(aes(fill = broadband_usage)) + 
  geom_sf(size = .05) +
    ## thickness of county borders
  scale_fill_gradient2(labels = percent, low = "red", high = "blue",
                       midpoint = .24) +
    ## change scale
  theme_map() +
  coord_sf() +
  labs(fill = "Broadband usage (%)")

## set midpoint based on distribution of broadband usage
bb_with_pop %>%
  ggplot(aes(broadband_usage)) +
  geom_histogram()

median(bb_with_pop$broadband_usage, na.rm = TRUE)

# reproducing the map from The Verge
counties_sf %>%
  inner_join(bb_with_pop, by = c(GEOID = "county_id")) %>%
  filter(!state %in% c("Alaska", "Hawaii")) %>%
  ggplot(aes(fill = broadband_usage >= .15 )) + 
  geom_sf(size = .05) +
  scale_fill_manual(values = c("red", "grey")) +
  theme_map() +
  coord_sf() +
  labs(fill = "Broadband usage (%)")

# of population
counties_sf %>%
  inner_join(bb_with_pop, by = c(GEOID = "county_id")) %>%
  filter(!state %in% c("Alaska", "Hawaii")) %>%
  ggplot(aes(fill = population_2019)) + 
  geom_sf(size = .05) +
    ## thickness of county borders
  scale_fill_gradient2(low = "red", high = "blue",
                       trans = "log10", ## transform scale to log10
                       midpoint = 4.5) +
  theme_map() +
  coord_sf() +
  labs(fill = "Population in 2019")


# let's look at median income
bbz_zip_join %>%
  filter(population > 5000) %>%
  ggplot(aes(median_household_income, broadband_usage)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  scale_x_log10()

```


Correlation by state:

```{r, include = TRUE}

## try geofacet package
library(geofacet)

bbz_zip_join %>%
  filter(population >= 10000) %>%
  ggplot(aes(median_household_income, broadband_usage)) +
  geom_point(size = 0.05, alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_geo(~ state) + 
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Median Household Income ($)",
       y = "Broadband usage in 2019")

bbz_zip_join %>%
  ggplot(aes(median_household_income)) +
  geom_histogram()

bbz_zip_join %>%
  filter(population >= 10000) %>%
  ggplot(aes(median_household_income, broadband_usage)) +
  geom_point(size = 0.05, alpha = 0.2) +
  geom_smooth(method = "lowess") +
  facet_geo(~ state) + 
  scale_x_log10(labels = comma_format()) +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Median Household Income (log10)",
       y = "Broadband usage in 2019")

```



```{r, include = TRUE} 
library(urbnmapr)

bb_with_pop %>%
  count(state, sort = T) %>%
  View()

# let's plot alaska and hawaii too
counties_sf <- get_urbn_map("counties", sf = TRUE)

broadband_availability_per_fcc

counties_sf %>%
  inner_join(bb_with_pop, by = c(county_fips = "county_id")) %>%
  ggplot() + 
  geom_sf(mapping = aes(fill = broadband_availability_per_fcc),
          color = "#ffffff", size = .1) +
  scale_fill_gradient2(labels = scales::percent,
                       midpoint = median(bb_with_pop$broadband_availability_per_fcc, na.rm = TRUE)) +
  theme_map() +
  coord_sf() +
  theme(legend.position="bottom",
        title = element_text(family = "Times"),
        plot.title = element_text(size = 26)) +
  labs(title = "Access to the internet in the U.S.",
       fill = "Broadband availability per fcc")

```

Adding fonts
```{r, include = TRUE} 

library(showtext)
font_add_google(name = "Montserrat", family = "montserrat")
  ## find google fonts: https://fonts.google.com/
font_add_google(name = "Codystar", family = "codystar")

showtext_auto()
  ## turn on showtext to load the new font


library(ggdark)

counties_sf %>%
  inner_join(bb_with_pop, by = c(county_fips = "county_id")) %>%
  ggplot() + 
  geom_sf(mapping = aes(fill = broadband_availability_per_fcc),
          color = "#ffffff", size = .1) +
  scale_fill_gradient2(labels = scales::percent,
                       midpoint = median(bb_with_pop$broadband_availability_per_fcc, na.rm = TRUE)) +
  dark_theme_minimal() +
  theme(plot.title = element_text(family = "codystar", size = 100, color = "white", hjust = 0.5),
        plot.subtitle = element_text(family = "montserrat", size = 12, color = "white", hjust = 0.5),
        plot.caption = element_text(family = "montserrat", size = 8, color = "white"),
        legend.position = "bottom",
        legend.text = element_text(family = "montserrat", size = 8, color = "white"),
        legend.title = element_text(family = "montserrat", size = 12, color = "white", hjust = 0.5),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  labs(title = "Access to \n the internet",
       subtitle = "\n Broadband availability is defined as the \n percent of people per county with access to \n fixed terrestrial broadband at speeds of \n 25 Mbps/3 Mbps as of the end of 2017",
       fill = "Broadband availability",
       caption = "Data: Microsoft | Plot: @nozominaka") 
  
ggsave(width = 18, height = 12, dpi = 320, filename = "InternetAccess_plot.png")

```

